Modelling complex population structure using $F$-statistics and Principal Component Analysis

### 2022-01-13 fstats_pca_lewontin.tex
bla

The most widely considered alternative model is an admixture graph \citep{patterson2012}, an example is given in Figure \ref{fig:admix}A. Here, the (typically unobserved) population $X_y$ is generated by a mixture of individuals from the ancestors of $X_2$ and $X_3$. Over time, genetic drift will change $X_y$ to $X_x$, which is the admixed population we observe. This will result in $F_4$-statistics that are non-zero, and, in some cases, in negative $F_3$-statistics \citep[exact conditions can be found in ][]{peter2016}.

>>>>

The most widely considered alternative model is an admixture graph \citep{patterson2012}, an example is given in Figure \ref{fig:admix}A. Here, the (typically unobserved) population $X_y$ is generated by a mixture of individuals from the ancestors of $X_2$ and $X_3$. Over time, genetic drift will change $X_y$ to $X_x$, which is the admixed population we observe. This will result in $F_4$-statistics that are non-zero, and, in some cases, in negative $F_3$-statistics \citep[exact conditions can be found in][]{peter2016}.

In many applications, it is useful to distinguish between $F$-statistics as parameters (e. g. length of a branch that is estimated) and an estimated $f$-statistic from data, denoted by convention by a lower case $f$ \citep{patterson2012}. Because PCA in population genetics is most commonly done on complete data without an estimation step, I do not make this distinction and use (upper-case) $F$-statistics throughout the manuscript.
